Interfering with Noninterference \iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF and...?)}}\fi

### 2020-01-17 main.tex
First stab at the (closure) properties of contoured hyperproperties. Too tedious?

A {\em contoured trace} is a finite or infinite sequence
$\seq{(\contour,\mach)} = (\contour_0,\mach_0),\,
(\contour_1,\mach_1),\ldots$ of (contour, machine state) pairs in which
adjacent machine states are related by $\stepstounderfull$---i.e., in which
$\mach_{i-1} \stepstounderfull \mach_i$ for each $i > 0$ in the sequence.

\subsection{Observations}

A {\em component observation} is either a word or the special value $\bot$.
A {\em state observation} is a map from machine components to component
observations.

An {\em observation} is a nonempty, finite or infinite sequence of state
observations in which (1) adjacent elements are always distinct (``no
stuttering'') and (2) any component $\component$ that is mapped to $\bot$ by
the the $j^{th}$ state observation in the sequence must also be mapped to
$\bot$ by all observations earlier than $j$ in the sequence (i.e., once a
component's value changes from $\bot$ to anything else, it never changes
back to $\bot$ again).

In particular, the observation $\obstrace(\seq{(\contour,\mach)})$ arising
from a contoured trace $\seq{(\contour,\mach)}$ is sequence of state
observations $\stateobs_0,\, \stateobs_1,\, \ldots$ defined informally as
follows:
\begin{itemize}
\item $\stateobs_0$ maps every visible component (where $\contour_0 = \VIS$)
to its value in $\mach_0$ and every hidden
component to $\bot$; 
\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for each
visible component (where $\contour_{i+1} = \VIS$), copying its
value from $\mach_{i+1}$, and (2) for each
hidden component, where $\contour_i = \VIS$, copying its
value from $\stateobs_{i+1}$, and (3) if the resulting $\stateobs_{i+1}$ is
identical to $\stateobs_{i}$, then skipping $\stateobs_{i+1}$.
\iftext\bcp{Too informal!!}\fi
\end{itemize}

@@

% An {\em observer} $\observer \in CONTOURS \rightarrow (\MA \exists X.\; (\CONTOURS \rightarrow X) \times
% (X \rightarrow \CONTOURS) \times (\MACHS \times X \rightharpoonup \MACHS
% \times X) $
% comprises
% \begin{enumerate*}
%   \item a private type $X$,
%   \item a function taking a contour to an initial value of $X$, and
%   \item a function that allows ``reading off'' a contour from an $X$, and
%   \item a partial step function of type
%   $(\MACHS \times X) \rightharpoonup X$.
% \end{enumerate*}

A {\em contoured hyperproperty} is a set of pairs of contoured
traces\iftext{} \bcp{Maybe prefix closed?}\fi.
% \bcp{Should we require that it contain all possible starting states and
%   contours (though of course the sequences can be singletons in many
%   cases!)?  What does it mean if it doesn't?}

@@

\leo{Low should be VIS here right?}
Given a machine state $\mach$ and a contour $\contour$, a {\em variant} of
$\mach$ with respect to $\contour$ is any machine state $\mach'$ whose
components agree with $\mach$ on all components marked $\LOW$ by
$\contour$---i.e., if $\contour(\component) = (\LOW$,\,\underscore), then
$\mach(\component) = \mach'(\component)$.

>>>>

A {\em contoured trace} of length $n$, where $n$ can be finite or
infinite, is a length-$n$ sequence $\seq{(\contour,\mach)}_n =
(\contour_0,\mach_0),\, (\contour_1,\mach_1),\ldots$ of (contour,
machine state) pairs in which adjacent machine states are related by
$\stepstounderfull$---i.e., in which $\mach_{i-1} \stepstounderfull
\mach_i$ for each $i > 0$ in the sequence.

% 
%\subsection{Observations}
% 
%A {\em component observation} is either a word or the special value $\bot$.
%A {\em state observation} is a map from machine components to component
%observations.
% 
%An {\em observation} is a nonempty, finite or infinite sequence of state
%observations in which (1) adjacent elements are always distinct (``no
%stuttering'') and (2) any component $\component$ that is mapped to $\bot$ by
%the the $j^{th}$ state observation in the sequence must also be mapped to
%$\bot$ by all observations earlier than $j$ in the sequence (i.e., once a
%component's value changes from $\bot$ to anything else, it never changes
%back to $\bot$ again).
% 
%In particular, the observation $\obstrace(\seq{(\contour,\mach)})$ arising
%from a contoured trace $\seq{(\contour,\mach)}$ is sequence of state
%observations $\stateobs_0,\, \stateobs_1,\, \ldots$ defined informally as
%follows:
%\begin{itemize}
%\item $\stateobs_0$ maps every visible component (where $\contour_0 = \VIS$)
%to its value in $\mach_0$ and every hidden
%component to $\bot$; 
%\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for each
%visible component (where $\contour_{i+1} = \VIS$), copying its
%value from $\mach_{i+1}$, and (2) for each
%hidden component, where $\contour_i = \VIS$, copying its
%value from $\stateobs_{i+1}$, and (3) if the resulting $\stateobs_{i+1}$ is
%identical to $\stateobs_{i}$, then skipping $\stateobs_{i+1}$.
%\iftext\bcp{Too informal!!}\fi
%\end{itemize}

@@

% An {\em observer} $\observer \in CONTOURS \rightarrow (\MA \exists X.\; (\CONTOURS \rightarrow X) \times
% (X \rightarrow \CONTOURS) \times (\MACHS \times X \rightharpoonup \MACHS
% \times X) $
% comprises
% \begin{enumerate*}
%   \item a private type $X$,
%   \item a function taking a contour to an initial value of $X$, and
%   \item a function that allows ``reading off'' a contour from an $X$, and
%   \item a partial step function of type
%   $(\MACHS \times X) \rightharpoonup X$.
% \end{enumerate*}
\leo{The following all assumes lockstep-ness:}

\leo{I think we need variants from here... Should we call it
  indistinguishability?}

Given a machine state $\mach$ and a contour $\contour$, a {\em
  variant} of $\mach$ with respect to $\contour$ is any machine state
$\mach'$ whose components agree with $\mach$ on all components marked
$\VIS$ by $\contour$---i.e., if $\contour(\component) = \VIS$, then
$\mach(\component) = \mach'(\component)$.

A {\em contoured hyperproperty} is a set of pairs of contoured
traces such that:
\begin{itemize}
\item {\em same-length:}

  For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C',
    M')}_m$ in the set, $n = m$.

\item {\em contour-lockstep:}

  For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C',
    M')}_n$ in the set, $C_i = C_i'$ for all $i<n$.

\item {\em prefix-closed:}

  For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C,
    M')}_n$ in the set, the pair $\overline{(C, M)}_m$ and
  $\overline{(C, M')}_m$ is also in the set if $m < n$.

\item {\em suffix-closed:}

  For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C,
    M')}_n$ in the set, and for every $0 \leq i < n$, there exists in
  the set a pair of traces $\overline{(C', N)}_m$ and $\overline{(C',
    N')}_m$ such that $C_0' = C_i$, $N_0 = M_i$, and $N_0' = M_0'$.

\item {\em variant-closed:}

  For every pair of traces $\overline{(C, M)}$ and $\overline{(C,
    M')}$ in the set, and for every machine state $N_0$ such that
  $N_0$ is a variant of $M_0$, there exists in the set a pair of
  traces $\overline{(C, M)}_m$ and $\overline{(C, N)}_m$.

\item {\em contour-wellformedness:}

  For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C,
    M')}_n$ in the set, $C_0 \leq C_i$ for all $i<n$, where $C \leq
  C'$ if $\forall \component, C(\component) \leq C'(\component)$ and
  $\VIS < \HID$.

\end{itemize}
\iftext{} \bcp{Maybe prefix closed?}\fi.
% \bcp{Should we require that it contain all possible starting states and
%   contours (though of course the sequences can be singletons in many
%   cases!)?  What does it mean if it doesn't?}

@@

Given a machine state $\mach$ and a contour $\contour$, a {\em variant} of
$\mach$ with respect to $\contour$ is any machine state $\mach'$ whose
components agree with $\mach$ on all components marked $\LOW$ by
$\contour$---i.e., if $\contour(\component) = (\LOW$,\,\underscore), then
$\mach(\component) = \mach'(\component)$.
